### 2025-03-18 code/eda-code/eda.qmd
Part 3

---
title: exploratory analysis script"
date: "2024-02-07"
output: html_document
---

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) 
library(dplyr)
library(skimr)
library(ggplot2)
```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

# Data exploration through tables

Showing a bit of code to produce and save a summary table.

```{r}
summary_df = skimr::skim(mydata)
print(summary_df)
# save to file
summarytable_file = here("results", "tables", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

We are saving the results to the `results` folder. Depending on how many tables/figures you have, it might make sense to have separate folders for each. And/or you could have separate folders for exploratory tables/figures and for final tables/figures. Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.

# Data exploration through figures

#Making a table for the clinical variable 
```{r}
table(processeddata$clinical)
```
#doing the same for food
```{r}
table(processeddata$food)
```
#Now lets visualize these tables using bar graphs
```{r}
ggplot(mydata, aes(x = clinical)) + 
  geom_bar(fill = "steelblue") + 
  labs(title = "Yes/No Distribution", x = "Response", y = "Count") + 
  theme_minimal()

```
```{r}
library(ggplot2)

ggplot(mydata, aes(x = food)) + 
  geom_bar(fill = "orange") + 
  labs(title = "Yes/No Distribution", x = "Response", y = "Count") + 
  theme_minimal()

```
#cross tabulation
```{r}
table(mydata$food, mydata$clinical)
```
#We can also look at some distribution across the country since we have some state data
```{r}

mydata %>%
  filter(food == "Yes") %>%  # Keep only "Yes" cases
  count(Location) %>%  # Count occurrences per location
  ggplot(aes(x = Location, y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") + 
  labs(title = "Count of food isolated listeria Cases Across Locations", 
       x = "Location", 
       y = "Count of Yes") + 
  theme_minimal() +
  theme(axis.text.x = element_text(size = 6,angle = 45, hjust = 1))  # Rotate x-axis labels for readability

```
#The left most bar is USA as a whole as some entries do not report the specific state they are from, the rest are states and territories. We see that NY has more cases than other states

#Some entries show the isolation source so lets make a boxplot showing the sources for some clinical isolates.
```{r}
library(ggplot2)
library(dplyr)

mydata %>%
  dplyr::filter(clinical == "Yes") %>%  
  count(`Isolation source`) %>%
  ggplot(aes(x = `Isolation source`, y = n)) +
  geom_boxplot(fill = "darkblue") +
  labs(title = "Count of Clinical 'Yes' Across Isolation Sources", 
       x = "Isolation Source", 
       y = "Count of Yes") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 4, angle = 45, hjust = 1))

#We see there are a ton of sources so we I will filter it to the top 10.
```

```{r}
mydata %>% 
  dplyr::filter(clinical == "Yes") %>%
  count(`Isolation source`) %>%
  top_n(10, n) %>%  # Keep only the top 10 sources
  ggplot(aes(x = reorder(`Isolation source`, -n), y = n)) +
  geom_bar(stat = "identity", fill = "darkblue") +
  labs(title = "Top 10 Clinical 'Yes' Isolation Sources", 
       x = "Isolation Source", 
       y = "Count of Yes") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1))

```
#Not sure how pork ended up in the clinical category...
## Now I will try the same thing on food products

```{r}
mydata %>% 
  dplyr::filter(food == "Yes") %>%
  count(`Isolation source`) %>%
  top_n(10, n) %>%  # Keep only the top 10 sources
  ggplot(aes(x = reorder(`Isolation source`, -n), y = n)) +
  geom_bar(stat = "identity", fill = "darkblue") +
  labs(title = "Top 10 Food 'Yes' Isolation Sources", 
       x = "Isolation Source", 
       y = "Count of Yes") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1))
```
#Seeing a few issues with my data cleaning, I will go back to processing file and fix it up

# Notes

For your own explorations, tables and figures can be "quick and dirty". As long as you can see what's going on, there is no need to polish them. That's in contrast to figures you'll produce for your final products (paper, report, presentation, website, etc.). Those should look as nice, polished and easy to understand as possible.