### 2025-03-20 code/analysis-code/statistical-analysis.qmd
feedback updates

---
title: 
author: 
date: 
output: html_document
---

#Here I will perform some basic statistical analysis.

#Lets start with a basic glm of time and food and time and clinical. We will also do some visualizations using a histrogram