### 2025-03-24 products/manuscript/Manuscript.qmd

---
title: "Exploring clinical and non-clinical(food) Listeria monocytogenes isolates "
author: "Alexis Gonzalez"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

**Authors**

-   Alexis Gonzalez $^{1,2,*}$ (ORCID: 0000-0000-1234-5678)

**Author affiliations**

1.  College of Public Health, University of Georgia, Athens, GA, USA.

{{< pagebreak >}}

# Part 1 Project Idea

For this project I will utilize data from the National Center for Biotechnology Information's Pathogen Detection system. This is a database that collects genomic data from a various different sources such as state laboratories, private laboratories, and government agencies like CDC, FDA, USDA, and EPA. It is primarily used a surveillance tool to monitor isolates,clusters, antimicrobial resistance, and active outbreaks. As someone that keeps track of foodborne pathogen outbreaks, I have noticed the increase in Listeria monocytogenes outbreaks. With the NCBI data I would like to visualize the occurence of new listeria strains, compare trees for clinical vs nonclincal isolates, and maybe see the correlation between the number of clinical isolates to the number of non-clincal isolates, maybe investigate how often non-clinical isolates have a clinical in the same cluster. I could also look at distribution across the United States. There is some data for other countries, but far less than the U.S. . For this project I will filter the data to show isolates collected from January 1, 2015 to January 1, 2025. This gives me 22,980 isolates and 2,365 clusters.

{{< pagebreak >}}

# Summary/Abstract

Will complete at the end

{{< pagebreak >}}

# Introduction

## General Background Information

*Listeria monocytogenes* is a ubiquitous, gram positive, intra-cellular pathogen renowned for its adaptability and resilience (cite). This bacterium can be found in a wide range of environments such as soil, water, animals, and vegetation. Due to its ability to proliferate in food products and low temperatures, it is considered a major food-borne illness. The ingestion of Listeria-infected food may lead to listeriosis, a gastrointestinal illness that causes vomiting, diarrhea, and even death. It is of even greater concern in pregnant individuals, as its forceful invasion tactics allow the bacterium to cross the placental barrier, infecting the fetus; this can lead to miscarriages and stillbirths\[cite\]. Listeriosis causes 1,600 infections and 260 deaths each year with an over 95% hospitalization rate. Listeria can be found in a wide range of meats, produce, and dairy, such as deli meats, soft cheese, and melons. The Centers for Disease Control and Prevention has deemed *Listeria monocytogenes* as a nationally notifiable disease which is a disease that federal agencies, healthcare providers, and laboratories are required to report. All positive isolate data found in the environment, food samples, clinical samples and animal samples are uploaded to the National Center for Biotechnology Information's Pathogen Detection system.

## Description of data and data source

This project will utilize two data sources

1.  National Center for Biotechnology Information's Pathogen Detection system
2.  CDC National Notifiable Disease Surveillance System

## Questions/Hypotheses to be addressed

1.  How has the presence of listeria in food products changes over time (NCBI data)
2.  How has the rate of listeriosis cases changed over time
3.  How does the positive rate of food cases compare to the rate of listeriosis cases
4.  Comparing and contrasting the isolates found in food to the isolates found in clinical samples.
5.  What are some commonalities between establishments and locations where listeria is found and where cases are diagnosed
6.  Follow up to 4, are these large scale companies or small scale companies.
7.  Maybe a prediction model based to predict listeriosis cases based on number of positive isolates found?

{{< pagebreak >}}

# Methods

## Schematic of workflow

```         
```

## Data import and cleaning

Before important the NCBI data, I filtered the data to what I wanted using the sites filters. I filtered data down to organism = Listeria monocytogenes, location = USA , this left me with 33,264 isolates.

I will briefly use the NNDSS data because unfortunately, it is only available from 2016 to 2022. The NCBI data contains clinical isolates that date farther back to we will consider those isolates to be "cases". Those cases are what we will compare to the source isolates (environmental, food, etc)

The data cleaning code can be found in the processingfile under data-\>rawdata. The cleaning processed involved removing variables that I would not be using such as biosample numbers, assembly numbers, and variables that contained no data. Next I made some changes to date entries. They originally contained the time the isolate was uploaded so I removed that. I also created 2 new variables showing if the isolate was from a food sample or a clinical sample . The variable "Clinical" has a yes or no entry for if the sample is clinical, the same goes for the Food variable.

## Statistical analysis

*Explain anything related to your statistical analyses.*

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

The full code for the EDA is under code-\>eda-code-\>eda.qmd

## Basic statistical analysis

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
```

## Full analysis

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

## Strengths and Limitations

## Conclusions

{{< pagebreak >}}

# References

```{r}

```