### 2025-04-06 code/analysis-code/statistical-analysis.qmd
adding some figures to manuscript

---
title: 
author: 
date: 
output: html_document
---
# loading in packages
```{r}
library(tidyverse)
library(here) 
library(dplyr)
library(skimr)
library(ggplot2)
library(tidymodels)
```
#Load in the data
```{r}
#Path to data. 
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

#Here I will perform some basic statistical analysis.

#Lets start with a basic glm of time and food and time and clinical. We will also do some visualizations using a histrogram

#To do this I have to convert my date variable into a numeric variable.

```{r}
# Load necessary package
library(lubridate)

"# Sample data
mydata <- data.frame(
  date `Create date` = as.Date,
  outcome = clinical  # Yes = 1, No = 0
)"

# Recode Yes = 1, No = 0
mydata$clinical_numeric <- ifelse(mydata$clinical == "Yes", 1, 0)

# Convert date to numeric (days since first observation)
mydata$days_since_start <- as.numeric(difftime(mydata$`Create date`, min(mydata$`Create date`), units = "days"))

# Logistic regression
model <- glm(clinical_numeric ~ days_since_start, data = mydata, family = binomial)

# Model summary
summary(model)

```
```{r}

ggplot(mydata, aes(x = days_since_start, y = clinical_numeric)) +
  geom_jitter(height = 0.05, width = 5, alpha = 0.5) +  
  geom_line(data = mydata, aes(x = days_since_start, y = predicted_prob), color = "blue", size = 1) +
  labs(title = "Logistic Regression: Probability of Clinical Isolates Over Time",
       x = "Days Since Start",
       y = "Probability of Outcome (Yes = 1)") +
  theme_minimal()

```
```{r}
p <- ggplot(mydata, aes(x = days_since_start, fill = factor(clinical_numeric))) +  # Fill by outcome (0 or 1)
  geom_histogram(position = "identity", 
                 alpha = 0.5, 
                 bins = 30, 
                 color = "black") +  # Standard histogram
  labs(title = "Histogram of clinical isolates Over Time",
       x = "Days Since Start",
       y = "Count of Outcome (0 or 1)",
       fill = "Outcome") +
  theme_minimal()

figure_file=here("results/figures","clinicalxtime.png")
ggsave(filename=figure_file, plot = p, width = 8, height=6)
```

```