### 2025-04-21 code/analysis-code/statistical-analysis.qmd
updates to manuscript

---
title: 
author: 
date: 
output: html_document
---
# loading in packages
```{r}
library(tidyverse)
library(here) 
library(dplyr)
library(skimr)
library(ggplot2)
library(tidymodels)
library(lubridate)
library(ranger)
library(parsnip)
library(poissonreg)
library(randomForest)
```
#Load in the data
```{r}
#Path to data. 
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

#Here I will perform some basic statistical analysis.

#Lets start with a basic glm of time and food and time and clinical. We will also do some visualizations using a histrogram

#To do this I have to convert my date variable into a numeric variable.

```{r}
# Load necessary package
library(lubridate)

"# Sample data
mydata <- data.frame(
  date `Create date` = as.Date,
  outcome = clinical  # Yes = 1, No = 0
)"

# Recode Yes = 1, No = 0
mydata$clinical_numeric <- ifelse(mydata$clinical == "Yes", 1, 0)

# Convert date to numeric (days since first observation)
mydata$days_since_start <- as.numeric(difftime(mydata$`Create date`, min(mydata$`Create date`), units = "days"))

# Logistic regression
model <- glm(clinical_numeric ~ days_since_start, data = mydata, family = binomial)

# Model summary
summary(model)

```

# temporal analysis of clinical isolates. Remember that every clinical isolate is like a human case of Listeriosis.
#For this analysis I will make some changes to my data by separating the time data into seasons.Winter is considered December, January and February; spring is March through May; summer is June through August; and fall or autumn is September through November.

```{r}
mydata <- mydata %>%
  drop_na()%>%
  mutate(
    `Create date` = as.Date(`Create date`),
    year = year(`Create date`),
    day_of_year= yday(`Create date`),
    week = week(`Create date`),
    month = month(`Create date`),
    season = case_when(
      month %in% c(12, 1, 2)  ~ "Winter",
      month %in% c(3, 4, 5)   ~ "Spring",
      month %in% c(6, 7, 8)   ~ "Summer",
      month %in% c(9, 10, 11) ~ "Fall"
    ),
    season = factor(season, levels = c("Winter", "Spring", "Summer", "Fall")),
    `Isolation source` = factor(`Isolation source`)
  )
```

#Feature selection
```{r}
model1_data <- mydata %>%
  select(`Isolation type`,year,month,day_of_year,week,season)
```

#Split the data
```{r}
set.seed(1234)
data_split <- initial_split(model1_data, prop = 0.8, strata = `Isolation type`)
train_data <-training(data_split)
test_data <-testing(data_split)
```
#Create recipe
```{r}
rf_recipe <- recipe(`Isolation type` ~ month + year + day_of_year + week + season, data = train_data) %>%
  step_dummy(all_nominal_predictors()) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_numeric_predictors())
```

#Define the model
```{r}
rf_mod <-rand_forest(trees = 150, min_n =5) %>%
  set_mode("classification") %>%
  set_engine("ranger",importance="impurity")
```
Create workflow
```{r}
workflow1 <- workflow() %>%
  add_model(rf_mod)%>%
  add_recipe(rf_recipe)
```
#fit the random forest model
```{r}
fit1 <- workflow1 %>%
  fit(data = train_data)
```
#Make predictions for evaluation
```{r}
predictions_rf <-fit1 %>%
  predict(new_data = test_data) %>%
  bind_cols(test_data)

```
```{r}
#set isolation source as a factor
predictions_rf <- predictions_rf %>%
  mutate(`Isolation type` = as.factor(`Isolation type`))

```

#Run metrics to evaluate the model
```{r}
# Create a metric set for classification
classification_metrics <- metric_set(accuracy, precision, recall, f_meas)
```

```{r}
#  use it on predictions
metrics_rf <- predictions_rf %>%
  classification_metrics(truth = `Isolation type`, estimate = .pred_class)

```

```{r}
metrics_rf <- predictions_rf %>%
  metrics(truth = `Isolation type`, estimate = .pred_class)
print(metrics_rf)
```
#Plot the results of this model

```{r}
library(vip)  # for variable importance plot

# Extract the fitted model from the workflow
rf_fit <- extract_fit_parsnip(fit1)

# Plot variable importance
vip(rf_fit$fit)

```

#Poisson modeling using month

```{r}
# Create monthly clinical counts
monthly_counts <- mydata %>%
  filter(clinical == "Yes") %>%
  mutate(month = floor_date(`Create date`, "month")) %>%
  group_by(month) %>%
  summarize(clinical_count = n()) %>%
  ungroup()
#filling in gaps with 0
monthly_counts <- monthly_counts %>%
  complete(month = seq.Date(min(month), max(month), by = "month"),
           fill = list(clinical_count = 0))
#Feature engineering, extracting the month from the data
monthly_counts <- monthly_counts %>%
  mutate(month_name = factor(month(month, label = TRUE), ordered = FALSE))

```

```{r}
library(tidymodels)
library(parsnip)
#Splitting the data 
set.seed(123)
data_split2 <- initial_time_split(monthly_counts, prop = 0.8)
train_data2 <- training(data_split2)
test_data2  <- testing(data_split2)

```
```{r}
#create a recipe 
poisson_recipe <- recipe(clinical_count ~ month_name, data = train_data2) %>%
  step_dummy(all_nominal_predictors())

```
```{r}
#Select the model
poisson_model <- 
  poisson_reg(
    mode = "regression",
    penalty = NULL,
    mixture = NULL,
    engine ="glm"
  )

```
```{r}
#Create a workflow 
library(poissonreg)
poisson_wf <-workflow() %>%
  add_model(poisson_model) %>%
  add_recipe(poisson_recipe)
```
```{r}
#fit the model
poisson_fit <- poisson_wf %>%
  fit(data = train_data2)
```
```{r}
#make predictions
predictions <- predict(poisson_fit, new_data = train_data2) %>%
  bind_cols(train_data2)
metrics(predictions, truth = clinical_count, estimate =.pred)
```
#Visualize results
```{r}
poisson_seasonal <- ggplot(predictions, aes(x = month)) +
  geom_line(aes(y = clinical_count), color = "black", linetype = "dashed") +
  geom_line(aes(y = .pred), color = "blue") +
  labs(title = "Predicted vs Actual Monthly Clinical Isolate Counts",
       y = "Clinical Count")
print(poisson_seasonal)
ggsave("/Users/alexisgonzalez/Desktop/MADA/Gonzalez-MADA-project/results/figures/poisson_seasonal.png", plot = poisson_seasonal, width = 8, height = 6, dpi = 300)
```
#Predict future isolate counts based on seasonality
```{r}
last_month <-max(monthly_counts$month)
```

```{r}
future_months <- tibble(
  month = seq.Date(from = last_month %m+% months(1),by = "month", length.out = 12)
)%>%
  mutate(month_name = factor(month(month, label = TRUE), ordered= FALSE))
```

```{r}
#make predictions
future_predictions <- predict(poisson_fit, new_data = future_months)%>%
  bind_cols(future_months)
```
```{r}
#visualize
# Combine historical + future for plotting
combined_months<- bind_rows(
  monthly_counts %>% select(month, clinical_count) %>% mutate(source = "Actual"),
  future_predictions %>% rename(clinical_count = .pred) %>% mutate(source = "Forecast")
)

# Plot
seasonalpreds <- ggplot(combined_data, aes(x = month, y = clinical_count, color = source)) +
  geom_line() +
  labs(title = "Actual and Forecasted Clinical Isolate Counts",
       y = "Clinical Count",
       x = "Month") +
  scale_color_manual(values = c("Actual" = "black", "Forecast" = "blue")) +
  theme_minimal()
print(seasonalpreds)
ggsave("/Users/alexisgonzalez/Desktop/MADA/Gonzalez-MADA-project/results/figures/seasonalpreds.png", plot = seasonalpreds, width = 8, height = 6, dpi = 300)
```

```{r}

# 1. Get the last month from your historical data
last_month <- max(monthly_counts$month)

# 2. Create the next 12 months
future_months <- tibble(
  month = seq.Date(from = last_month %m+% months(1), by = "month", length.out = 12)
) %>%
  mutate(month_name = factor(month(month, label = TRUE), ordered = FALSE))

# 3. Make predictions using your Poisson model (or other model)
future_predictions <- predict(poisson_fit, new_data = future_months) %>%
  bind_cols(future_months)

# 4. Combine historical and forecast data
combined_months <- bind_rows(
  monthly_counts %>% 
    select(month, clinical_count) %>% 
    mutate(source = "Actual"),
  future_predictions %>% 
    rename(clinical_count = .pred) %>% 
    mutate(source = "Forecast")
)

# 5. Plot
seasonalpreds <- ggplot(combined_months, aes(x = month, y = clinical_count, color = source)) +
  geom_line(size = 1.2) +
  labs(
    title = "Actual and Forecasted Clinical Isolate Counts",
    y = "Clinical Count",
    x = "Month"
  ) +
  scale_color_manual(values = c("Actual" = "black", "Forecast" = "blue")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# 6. Show plot
print(seasonalpreds)

# 7. Save to file
ggsave(
  "/Users/alexisgonzalez/Desktop/MADA/Gonzalez-MADA-project/results/figures/seasonalpreds.png", 
  plot = seasonalpreds, width = 8, height = 6

```

```{r}
library(mgcv)
gam_model <- gam(clinical_count ~ s(day_of_year), family = poisson(), data = daily_counts)
plot(gam_model)
ggsave("/Users/alexisgonzalez/Desktop/MADA/Gonzalez-MADA-project/results/figures/gam_model.png", plot = gam_model, width = 8, height = 6, dpi = 300)
```

```{r}
# Load necessary library
library(mgcv)
library(gratia) # for tidy ggplot GAM visualizations

# Example: Create ggplot-style plot from your GAM model
gam_plot <- draw(gam_model)  # `gam_model` is your fitted model

# Now save it
ggsave("/Users/alexisgonzalez/Desktop/MADA/Gonzalez-MADA-project/results/figures/gam_model.png",
       plot = gam_plot, width = 8, height = 6, dpi = 300)

```

```{r}

#  Split the data into training and testing sets
set.seed(123)
data_split <- initial_split(monthly_counts, prop = 0.8)
train_data <- training(data_split)
test_data <- testing(data_split)

#  Define the recipe (clinical_count predicted by food_count)
poisson_recipe <- recipe(clinical_count ~ food_count, data = train_data)

# Define the Poisson regression model
poisson_model <- poisson_reg(mode = "regression", engine = "glm")

#  Create a workflow 
poisson_wf <- workflow() %>%
  add_recipe(poisson_recipe) %>%
  add_model(poisson_model)

#  Fit the model
poisson_fit <- poisson_wf %>%
  fit(data = train_data)

# Make predictions on the training data
predictions <- predict(poisson_fit, new_data = train_data) %>%
  bind_cols(train_data)

#  Evaluate the model 
isolate_metrics <- metrics(predictions, truth = clinical_count, estimate = .pred)
print(isolate_metrics)

# Visualize the results
ggplot(predictions, aes(x = month)) +
  geom_col(aes(y = clinical_count), fill = "lightgray", alpha = 0.8) +
  geom_line(aes(y = .pred), color = "blue", size = 1) +
  labs(
    title = "Poisson Regression: Predicting Clinical Isolates from Food Isolates",
    y = "Clinical Isolate Count",
    x = "Month"
  ) +
  theme_minimal()
```
#Chart showing likelihood
```{r}

```

#Now I want to do one more advanced statistical analysis. This analysis will utilize the counts of clinical and counts of food isolates each year and see if the amount of clinical isolates can be predicted by the amount of food isolates. This purpose of this analysis is to see if federal reguulations are doing a good job of preventing illness.

#put my food and clinical variable into counts and group by year

```{r}
monthly_counts2 <- mydata %>%
  mutate(month = floor_date(`Create date`, "month")) %>%
  count(month, isolation_source) %>%
  tidyr::pivot_wider(names_from = isolation_source, values_from = n, values_fill = 0)
```

```{r}
set.seed(123)  # for reproducibility

rf_model <- randomForest(
  clinical ~ food + environmental,
  data = monthly_counts2,
  ntree = 500,          # number of trees
  importance = TRUE     # track variable importance
)
print(rf_model)

```

```{r}
#model_predict_clinical <- lm(clinical ~ food + environmental, data = monthly_counts2)
#summary(model)

```{r}

```

`
#Visualzing the future
```{r}
# Suppose you created this:
future_data <- data.frame(
  month = as.Date(c("2025-02-01","2025-03-01","2025-04-01","2025-05-01", "2025-06-01", "2025-07-01","2025-08-01","2025-09-01","2025-10-01","2025-11-01","2025-12-01")),
  food = c(300,220, 180, 250,140,122,116,235,96,175,225),
  environmental = c(350,150, 170, 200,220,140,267,178,155,200,280)
)
```

```{r}
# Predict clinical isolates
future_data$Predicted_Clinical <- predict(rf_model, newdata = future_data)
```

```{r}
# Predict clinical isolates for future months
future_data$Predicted_Clinical <- predict(rf_model, newdata = future_data)

# Combine with past data (no predicted clinicals in the past, so we set it to NA)
combined_plot_data <- monthly_counts2 %>%
  select(month, clinical) %>%
  mutate(Predicted_Clinical = NA,
         Source = "Historical") %>%
  bind_rows(
    future_data %>%
      transmute(month, clinical = NA, Predicted_Clinical, Source = "Forecast")
  )

# Filter combined_plot_data for the last 3 years
recent_data <- combined_plot_data %>%
  filter(month >= (max(month) %m-% years(3)))

# Plot
ggplot(recent_data, aes(x = month)) +
  geom_line(aes(y = Predicted_Clinical, color = Source), size = 1.2, linetype = "dashed") +
  geom_line(data = recent_data %>% filter(!is.na(clinical)),
            aes(y = clinical, color = "Observed"), size = 1) +
  scale_color_manual(values = c("Observed" = "steelblue", "Historical" = "tomato", "Forecast" = "forestgreen")) +
  labs(title = "Observed and Forecasted Clinical Isolates (Past 3 Years)",
       x = "Month",
       y = "Count",
       color = "Legend") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
```{r}
# Create new data (adjust as needed)
new_data <- data.frame(food = c(200), environmental = c(150))

# Predict clinical isolates
predict(model_predict_clinical, newdata = new_data)

```
#Machine Learning since this was just a simple linear regression
#Random forest
```{r}
#set seed and split data
set.seed(123)
data_split3 <- initial_split(model_data, prop = 0.8)
train_data3 <- training(data_split3)
test_data3 <- testing(data_split3)

```