### 2025-05-02 code/eda-code/eda.qmd
additions

---
title: exploratory analysis script"
date: "2024-02-07"
output: html_document
---

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) 
library(dplyr)
library(skimr)
library(ggplot2)
```

#Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```
#Creating an additional variable called "Isolation_source" This is different from the `Isolation Source` variable in that it will include food and not just list it under environmental/other
```{r}
mydata <- mydata %>%
  mutate(
    clinical = as.character(clinical),
    food = as.character(food),
    isolation_source = case_when(
      clinical == "Yes" ~ "clinical",
      food == "Yes" ~ "food",
      TRUE ~ "environmental"
    )
  )

```

# Data exploration through tables

#Making a table for the clinical variable 
```{r}
table(mydata$clinical)
```
#doing the same for food
```{r}
table(mydata$food)
```
```{r}
table(mydata$isolation_source)
```

#Table 1: counts across the 15 year period with range and mean

```{r}
library(kableExtra)

# Create new year and month columns
mydata <- mydata %>%
  mutate(
    year = year(`Create date`), 
    month = month(`Create date`, label = TRUE, abbr = TRUE)
  )

# Count records
counts <- mydata %>%
  group_by(year, month, isolation_source) %>%
  summarize(count = n(), .groups = "drop")

# Summarize by year and source
yearly_totals <- counts %>%
  group_by(isolation_source, year) %>%
  summarize(
    total = sum(count),
    highest_month = month[which.max(count)],
    lowest_month = month[which.min(count)],
    .groups = "drop"
  )

# Pivot for final table
table_data <- yearly_totals %>%
  pivot_wider(
    names_from = isolation_source,
    values_from = c(total, highest_month, lowest_month),
    names_sep = "_"
  ) %>%
  arrange(year)

# Create the styled table
table_output <- table_data %>%
  kbl(
    caption = "Yearly Isolate Summary by Source",
    col.names = c(
      "Year", 
      "Total Clinical", "Total Environmental", "Total Food", 
      "Highest Month (Clinical)", "Highest Month (Environmental)", "Highest Month (Food)", 
      "Lowest Month (Clinical)", "Lowest Month (Environmental)", "Lowest Month (Food)"
    )
  ) %>%
  kable_styling(
    full_width = FALSE, 
    position = "center", 
    bootstrap_options = c("striped", "hover", "condensed")
  )
```

```{r}
# Define the file path
summarytable_file <- here::here("results", "tables", "summarytable.rds")

# Save the styled table as an RDS file
saveRDS(table_output, file = summarytable_file)

```

#Now lets visualize these tables using bar graphs of counts for food clincial and env
```{r}

isolatetypes <- ggplot(mydata, aes(x = isolation_source)) +
  geom_bar(fill = "seagreen") +
  labs(title = "Isolation Source Distribution",
       x = "Isolation Source",
       y = "Count") +
  theme_minimal()
 print(isolatetypes)  
ggsave("/Users/alexisgonzalez/Desktop/MADA/Gonzalez-MADA-project/results/figures/isolatetypes.png", plot = isolatetypes, width = 8, height = 6, dpi = 300)
```
#As we see here, clinical cases are rare but that does not take away from the fact that they are usually severe.
```{r}
#cross tabulation
table(mydata$food, mydata$clinical)
```
#We can also look at some distribution across the country since we have some state data

```{r}

locations <- ggplot(
  mydata %>% 
    filter(food == "Yes", Location != "USA"),
  aes(x = Location)
) + 
  geom_bar(fill = "steelblue") + 
  labs(
    title = "Count of Food-Isolated Listeria Cases Across Locations",
    x = "Location",
    y = "Count of Isolates"
  ) + 
  theme_minimal() + 
  theme(axis.text.x = element_text(size = 6, angle = 45, hjust = 1))
print(locations)
ggsave("/Users/alexisgonzalez/Desktop/MADA/Gonzalez-MADA-project/results/figures/locations.png", plot = locations, width = 8, height = 6, dpi = 300)
```

```{r}
locations2 <- ggplot(mydata %>% filter(clinical == "Yes"), aes(x = Location)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Count of Clincally-Isolated Listeria Cases Across Locations",
       x = "Location",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 6, angle = 45, hjust = 1))
print(locations2)
```
#This only tells me that all cases of listeriosis were in the U.S. which i specified for so we wont use this graph.

#The left most bar is USA as a whole as some entries do not report the specific state they are from, the rest are states and territories. We see that NY has more cases than other states
```{r}
sourcestate <- mydata %>%
  filter(clinical == "Yes", !is.na(`Food origin`)) %>%  
  count(`Food origin`) %>%  
  ggplot(aes(x = `Food origin`, y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Count of Clinically Isolated Listeria Sources by Food Origin",
       x = "Food Origin",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 6, angle = 45, hjust = 1))

print(sourcestate)
```

```{r}
ggsave("/Users/alexisgonzalez/Desktop/MADA/Gonzalez-MADA-project/results/figures/sourcestate.png", plot = sourcestate, width = 8, height = 6, dpi = 300)
```
#Based on this graph we see that many of the clinical cases are coming from foods produced in Sweden...interesting. maybe we can use food origin as a predictor as well.

#The left most bar is USA as a whole as some entries do not report the specific state they are from, the rest are states and territories. We see that NY has more cases than other states

#Some entries show the isolation source so lets make a boxplot showing the sources for some clinical isolates.
```{r}
library(ggplot2)
library(dplyr)

mydata %>%
  dplyr::filter(clinical == "Yes") %>%  
  count(`Isolation source`) %>%
  ggplot(aes(x = `Isolation source`, y = n)) +
  geom_boxplot(fill = "darkblue") +
  labs(title = "Count of Clinical Isolates Across Isolation Sources", 
       x = "Isolation Source", 
       y = "Count of isolates") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 4, angle = 45, hjust = 1))

#We see there are a ton of sources so we I will filter it to the top 10.
```

```{r}
mydata %>% 
  dplyr::filter(clinical == "Yes") %>%
  count(`Isolation source`) %>%
  top_n(10, n) %>%  # Keep only the top 10 sources
  ggplot(aes(x = reorder(`Isolation source`, -n), y = n)) +
  geom_bar(stat = "identity", fill = "darkblue") +
  labs(title = "Top 10 Clinical 'Yes' Isolation Sources", 
       x = "Isolation Source", 
       y = "Count of isolates") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1))

```
#Not sure how pork ended up in the clinical category...
## Now I will try the same thing on food products

```{r}
mydata %>% 
  dplyr::filter(food == "Yes") %>%
  count(`Isolation source`) %>%
  top_n(10, n) %>%  # Keep only the top 10 sources
  ggplot(aes(x = reorder(`Isolation source`, -n), y = n)) +
  geom_bar(stat = "identity", fill = "darkblue") +
  labs(title = "Top 10 Food Isolation Sources", 
       x = "Isolation Source", 
       y = "Count of Yes") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1))
```
#We see here that the most food isolates for listeria monocytogenes are found in Ready To Eat products, that is a term used by Food Safety and Inspection Service to describe any food products that do not require cooking or sterilization to safely consume.

#plot of food isolates and clinical isolates across 10 years
```{r}
library(tidyverse)
library(lubridate)

#Prepare the data into monthly counts
monthly_counts <- mydata %>%
  mutate(month = floor_date(`Create date`, "month")) %>%
  group_by(month) %>%
  summarize(
    food_count = sum(food == "Yes"),
    clinical_count = sum(clinical == "Yes"),
    environmental_count = sum(isolation_source =="environmental")
  ) %>%
  ungroup()

#Convert to long format for plotting
monthly_long <- monthly_counts %>%
  pivot_longer(cols = c(food_count, clinical_count, environmental_count),
               names_to = "source_type",
               values_to = "count")

#Plot monthly counts
ggplot(monthly_long, aes(x = month, y = count, color = source_type)) +
  geom_line(size = 1.2) +
  labs(
    title = "Monthly Salmonella Isolate Counts by Source",
    x = "Month",
    y = "Isolate Count",
    color = "Source Type"
  ) +
  scale_color_manual(values = c("environmental_count" = "forestgreen", "food_count" = "firebrick", "clinical_count"= "darkblue")) +
  theme_minimal()
```

```{r}
Isolatecounts <- ggplot(monthly_long, aes(x = month, y = count, fill = source_type)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(
    title = "Monthly Salmonella Isolate Counts by Source (Stacked Bars)",
    x = "Month",
    y = "Isolate Count",
    fill = "Source Type"
  ) +
  scale_fill_manual(values = c("environmental_count" = "forestgreen", "food_count" = "firebrick", "clinical_count"= "darkblue")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("/Users/alexisgonzalez/Desktop/MADA/Gonzalez-MADA-project/results/figures/isolatecounts.png", plot = Isolatecounts, width = 8, height = 6, dpi = 300)
print(Isolatecounts)
```