### 2025-05-02 code/processing-code/processingfile-v1.qmd

---
title: " cleaning script"
author: "Alexis Gonzalez"
date: "2024-02-07"
output: html_document
---

# Setup

Load needed packages. make sure they are installed.

```{r}
library(readxl) 
library(dplyr) 
library(tidyr)
library(skimr) 
library(here) 
library(lubridate)
```

# Data loading
```{r}
# path to data
# note the use of the here() package and not absolute paths
data_location <- here::here("data","raw-data","lmonoisolates.xls")
rawdata <- readxl::read_excel(data_location)
```

# Check data

```{r}
summary(rawdata)
dplyr::glimpse(rawdata)
summary(rawdata)
head(rawdata)
skimr::skim(rawdata)
```

# Cleaning

```{r}
lmono2 <- rawdata %>% 
  select(-c(Serovar, Assembly, BioSample))
```
Next I want to reformat the dates so that it removes the timestamp. I will just be keeping the year, month and day
```{r}
lmono2$`Create date` <- as.Date(lmono2$`Create date`)
print(lmono2$`Create date`)
```

```{r}
lmono2 <- lmono2 %>%
  mutate(
    year = year(`Create date`),
    month = month(`Create date`, label = TRUE, abbr = TRUE)
  ) %>%
  filter(year != 2011)

```
Next I will see the range of dates and select a 15 year time frame.
```{r}
lmono3 <- lmono2 %>%
  filter(`Create date` >= as.Date("2010-01-01") & `Create date` <= as.Date("2024-12-31"))
```
Finally I will check for any missing entries
```{r}
any(is.na(lmono3))
lmono3[!complete.cases(lmono3), ]  # Shows me what rows have missing data

```
#So this showed me that I need to remove another variable, all of the values for computed type and ...18 are NA!
```{r}
lmono4 <- lmono3 %>%
  select(-c(`Computed types`,...18))
```
#Now lets see where the missing data is 
```{r}
lmono4[!complete.cases(lmono4), ] 
```
#Now, I really want to compare food vs clinical isolates so I will make 2 new variables, One variable called Food amd one called Clinical. Food will have all entries that are food sources
```{r}
#first I am going to look at the entries, some entries are specific like hotdog or blood and some just say food or clinical, I want to include those specific ones as well so I need to see what I am working with. 
unique(lmono4$`Isolation type`)
#Some of the products under environmental other are food products so I will put environmnetal other into its own category and then filter out the non-food products since it is mostly foods.
```

```{r}
#This will create new variable "clinical"
lmono5 <- lmono4 %>%
  mutate(clinical = ifelse(`Isolation type` == "clinical", "Yes", "No"))
```

```{r}
lmono6 <- lmono5 %>%
  mutate(food = case_when(
    `Isolation type` %in% c("environmental/other", "Not Sliced-Pork", "Sliced-Pork", 
                            "Not Sliced-Turkey", "Sausage Products-Pork", 
                            "Salad/Spread/Pate-Pork", "Salad/Spread/Pate-Chicken", 
                            "Meat/Nonmeat Combination-Beef", "Meat/Nonmeat Combination", 
                            "Sliced-Beef", "Sliced-Turkey", "Not Sliced-Chicken", 
                            "Meat/Nonmeat Combination-Other Poultry", 
                            "Meat/Nonmeat Combination-Combination species", "raw", 
                            "Meat/Nonmeat Combination-Chicken", 
                            "Sausage Products-Unknown species", 
                            "Meat/Nonmeat Combination-Turkey", 
                            "Sausage Products-Beef") & 
    !`Isolation source` %in% c("environmental swab", "environmental swab sponge floor", 
                               "condensate tray", "squeegee", "chain", "wrench", 
                               "underhopper", "environmental swab sponge non food contact surface", 
                               "environmental swab sponge pallet in cooler", 
                               "environmental swab of shelf", 
                               "Environmental Isolate, sponge swab","environmental swab sponge","Environmental-factory","Environmental-Wheels on trolley for spinner basket","Environmental-Drain basin between retort 1 and 2","bottom of sliding door in pasta room to blast","bottom of sliding door in pasta room to blast","Environmental-Bottom of sliding door in pasta room to blast","	
environmental swab from drain","environmental", "water sediment","environmental: non-food-contact surface", "soil","manure", "Environmental: food-contact surface","environment","swab") ~ "Yes",
    TRUE ~ "No"
  ))

```
#Finally, I want to remove the observations that are stool samples that will be bovine, pigeon and catbird
```{r}
lmono_final <- lmono6 %>%
  filter(!`Isolation type` %in% c("Bos taurus\"\"", "pigeon", "catbird"))
```

#Creating an additional variable called "Isolation_source" This is different from the `Isolation Source` variable in that it will include food and not just list it under environmental/other
```{r}
lmono_final <- lmono_final %>%
  mutate(
    clinical = as.character(clinical),
    food = as.character(food),
    isolation_source = case_when(
      clinical == "Yes" ~ "clinical",
      food == "Yes" ~ "food",
      TRUE ~ "environmental"
    )
  )

```

```{r}
processeddata <-lmono_final
```

# Save data 
```{r}
save_data_location <- here::here("data","processed-data","processeddata.rds")
saveRDS(processeddata, file = save_data_location)
```